Meta-Analysis of the 'Ironic' Effects of Intergroup Contact

### 2022-02-17 manuscript/manuscript.Rmd
start revision
revision

---
title: "Meta-Analysis of the 'Ironic' Effects of Intergroup Contact"
header: "'IRONIC' EFFECTS META-ANALYSIS"
keywords: intergroup contact, meta-analysis, perceived discrimination, collective action, policy support
bibliography: references.bib
csl: apa.csl
nocite: | 
  @19, @45, @46, @91, @93, @244, @284, @286, @303, @322, @324, @325, @336, @401, @414, @423, @480, @609, @703, @733, @783, @789, @803, @813, @823, @829, @830, @856, @893, @894, @933, @956, @1020, @1042, @1045, @1114, @1163, @1221, @1248, @1376, @1386, @1395, @1407, @1411, @1412, @1494, @1525, @1549, @1576, @1614, @1658, @1695, @1743, @1762, @1807, @1850, @1891, @1914, @1949, @1966, @1993, @2001, @2023, @2033, @2075, @2257, @2309, @2333, @2341, @2375, @2380, @2381, @2382, @2383, @2385, @2391, @2392, @2394, @2395, @2396, @2397, @2398, @2399, @3000, @3005, @3051, @3054, @3060, @3134, @3196, @3201, @3205, @4000, @4001, @4002, @4003, @4004, @4005
abstract: 
  Growing evidence suggests that intergroup contact, psychology’s most-researched paradigm for reducing prejudice, has the ‘ironic’ effect of reducing support for social change in disadvantaged groups. We conducted a preregistered meta-analytic test of this effect across `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(dplyr::n_distinct(id)))` studies with `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(length(sample)))` samples of `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(sum(n)))` disadvantaged-group members. As predicted, intergroup contact was, on average, associated with less perceived injustice (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "pi"]), digits = 2), "$"))`), collective action (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ca"]), digits = 2), "$"))`), and support for reparative policies (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ps"]), digits = 2), "$"))`). However, these associations were small, variable, and consistent with alternative explanations. Across outcomes, `r with(readRDS("../results/results_for_abstract.rds"), paste0(round(min(p_pred)*100), "--", round(max(p_pred)*100), "\\%"))` of studies found *positive* associations with intergroup contact. Moderator analyses explained about a third of the between-sample variance, showing that, on average, associations were negative only in studies of adults that measured intergroup contact directly and were strongest in studies that examined short-term migration or (post-)colonial intergroup relations. We also found evidence for an alternative explanation for the apparent ‘ironic’ effects of intergroup contact as, after controlling for the positive association of negative contact with support for social change, positive contact was no longer associated with any of the outcomes. We close by discussing strengths and limitations of the available evidence and by highlighting open questions about the relationship between intergroup contact and support for social change in disadvantaged groups.
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: apa-template.tex
---

@@

Collective action is any action directed at improving the conditions of the disadvantaged ingroup [@wright_responding_1990]. Collective action is crucial to achieving social change because advantaged-group members rarely give up their advantages without sustained pressure from disadvantaged-group members [@blumer_race_1958; for a review, see @dixon_beyond_2012]. Researchers have argued that intergroup contact could discourage disadvantaged-group members from engaging in collective action by diminishing their perceived discrimination [@1743], by blurring boundaries between the disadvantaged ingroup and the advantaged outgroup [@813], by reducing negative attitudes toward the advantaged outgroup [@wright_struggle_2009], and by quelling their anger  [@2309; @2375]. To the extent that it discourages collective action, intergroup contact reduces support for social change.

@@

While narrative reviews of the growing literature on this topic have been published [@hassler_intergroup_2021; @macinnis_extending_2019; @mckeown_contact_2017], we present the first systematic review and meta-analysis of the relationship between intergroup contact and support for social change in disadvantaged groups. To test the hypothesized 'ironic' effects of intergroup contact, we evaluate the evidence that contact with advantaged-group members reduces disadvantaged-group members' perceived injustice, collective action, and support for reparative policies.

Our first objective was to systematically review the available evidence and to evaluate its strengths and limitations. Our second objective was to synthesize the available evidence to establish the direction and magnitude of the relationships between intergroup contact and the three outcomes. Doing so allowed us to determine whether the available evidence supports a negative or positive [see @pettigrew_recent_2011] relationship between contact and support for social change and to estimate the strength of this relationship. Our third objective was to estimate and explain the variability of this relationship across study settings, designs, and other potential moderators. For example, we examined whether any effects were confined to Western, Educated, Industrialized, Rich, Democratic (WEIRD) countries, the cultural context of most psychological research [@henrich_weirdest_2010]. Our fourth objective was to assess how robust the available evidence was to publication bias. Our final objective was to explore the available evidence for negative contact and ingroup contact as alternative explanations of the apparent 'ironic' effects of intergroup contact.

@@

We prepared a protocol in accordance with the PRISMA-P statement [@moher_preferred_2015] and preregistered it on the Open Science Framework.^[[https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0](https://osf.io/ryuev/?view_only=8938ac6178914da98bdda1da4a6e27f0)] Preregistration precludes undisclosed flexibility in study selection, outcome selection, and other decisions that influence effect size estimates—and thus prevents confirmation bias from determining the conclusions of a meta-analysis [@lakens_reproducibility_2016]. We report deviations from our protocol in the Supplemental Online Material (SOM). We make data, analysis scripts, and the fully reproducible manuscript available online.^[[https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a](https://osf.io/w5tqv/?view_only=5cbf7ad5b2b449a7a0f4ada0337b802a)]

@@

As preregistered, we searched titles, abstract, and keywords for relevant terms in four electronic databases. We searched for relevant articles in the *Scopus* and *PsycINFO* databases. We searched for gray literature in the *ProQuest Dissertations and Theses* database. We used similar non-exclusive search terms for all databases (see SOM). We searched databases on April 1, 2019 and again on April 1, 2020. We exported records and relevant metadata from each database. We removed duplicates using the *revtools* package [@westgate_revtools:_2019].

@@

### Outcomes Selection

@@

***Study Setting.*** For each sample, we recorded in what country (or countries) the data was collected ($\kappa = .97$) and what the disadvantaged ingroup and the advantaged outgroup were. We categorized each sample’s setting according to whether the source of the groups’ relative inequality was long-term migration (e.g., Asian Americans), short-term  migration (e.g., international students), slavery (e.g., Black Americans), colonization (e.g., Māori, Black South Africans), religion, caste, sexuality, or another distinction ($\kappa = .85$).

***Study Design.*** We categorized each study as either observational and cross-sectional, observational and longitudinal, quasi-experimental, experimental, an intervention, or other ($\kappa = .85$). We categorized each sample as either a student convenience sample, a non-student convenience sample, a probability or representative sample, or another kind of sample ($\kappa = .73$). We categorized the age group(s) in each sample as children ($\leq 12$ years), adolescents (13–18 years), or adults ($\geq 18$ years; $\kappa = .64$).

@@

***Predictor Variables.*** For each measure of intergroup contact, we coded whether it assessed contact with the specific advantaged outgroup of interest directly—or whether it assessed contact indirectly by, for example, asking what proportion of someone’s friends were not from the participants’ ingroup or what proportion of residents in someone’s neighbourhood were from the relevant outgroup ($\kappa = .82$).

@@

We estimated effect sizes using Bayesian random-effects meta-analysis models in *RStan* [@stan_development_team_rstan:_2020] which modeled the *z*-transformed correlation coefficients with a normal likelihood function: \begin{align*} z_{ij} &\sim \text{Normal}(\theta_{ij}, \sigma_{ij}) \\ \theta_{ij} &= \begin{cases} \mu + \beta_j\tau_J & \text{if } I_j = 1 \\ \mu + \beta_j\tau_J + \beta_i\tau_I & \text{if } I_j > 1 \end{cases} \\ \end{align*} where $z_{ij}$ is the observed effect size in sample $i$ of study $j$, $\sigma_{ij}$ is the sample standard error, and $\theta_{ij}$ is the estimated effect size. We estimated $z_{ij}$ as a function of the mean effect size $\mu$ and of two varying (random) intercepts, $\beta_i$ and $\beta_j$, with the corresponding standard deviations, $\tau_I$ and $\tau_J$. We used the non-centered parameterization to model the random effects. For studies that contained only one sample ($I_j = 1$), we only included $\beta_j\tau_J$, the study-wise deviation from the mean effect size $\mu$. For studies that contained more than one sample ($I_j > 1$), we also estimated $\beta_i\tau_I$, the sample-wise deviation from the study-specific effect size.

@@

Bayesian inference involves choosing a likelihood function and prior distributions. A likelihood function links the observed data to one or more model parameters and states how likely the observed data are given different values of said model parameters. Prior distributions state how plausible different values of said model parameters are before considering the observed data. Bayesian inference applies Bayes’ theorem to update prior distributions in light of the observed data to produce posterior distributions. In contrast to *p*-values and confidence intervals, the resulting posterior distributions have a straightforward interpretation as stating how plausible different values of the model parameters are given the observed data. We report point estimates, based on the median of posterior samples, and uncertainty intervals, based on the quantiles of posterior samples, that enclose the 95% most plausible estimates. In addition, we report the posterior probability, based on the proportion of posterior samples below zero, that the mean effect size is negative.

@@

@@

  # Leave out most influential studies
  d_loo <- read_rds("../results/results_robustness_checks_loo.rds") %>% 
    group_by(y_var, id) %>%
    median_qi(e_loo) %>% 
    top_n(n = 1, wt = abs(e_loo)) %>% 
    mutate(
      across(c(e_loo, .lower, .upper), ~f_num(., digits = 2)),
      text = paste0("$\\Delta r = ", e_loo, ", [", .lower, ", ", .upper, "]$")
    )

@@

As preregistered, we conducted two kinds of robustness checks. First, we assessed to what extent our findings were sensitive to choosing narrower, $\mu \sim \text{Normal}(0, 0.1)$, or wider, $\mu \sim \text{Normal}(0, 1)$, prior distributions. Choosing narrower or wider prior distribution did not affect mean effect size estimates for perceived injustice (`r d_priors$text[d_priors$y_var == "pi" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "pi" & d_priors$prior == "N(0, 1.0)"]`), collective action (`r d_priors$text[d_priors$y_var == "ca" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "ca" & d_priors$prior == "N(0, 1.0)"]`), and policy support (`r d_priors$text[d_priors$y_var == "ps" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "ps" & d_priors$prior == "N(0, 1.0)"]`). Second, we assessed to what extent our findings were sensitive to including or excluding influential studies by repeating the preregistered analyses $J$ times while leaving out one of $J$ studies each time and by calculating the mean absolute difference (*MAD*) for the estimated mean effect size across left-out studies. For perceived injustice (`r mad_loo$text[mad_loo$y_var == "pi"]`), collective action (`r mad_loo$text[mad_loo$y_var == "ca"]`), and policy support (`r mad_loo$text[mad_loo$y_var == "ps"]`), the *MAD* was small. Leaving out the most influential study, for example, did not change estimates of the mean effect size for the three outcomes (`r d_loo$text[d_loo$y_var == "pi"]`; `r d_loo$text[d_loo$y_var == "ca"]`; `r d_loo$text[d_loo$y_var == "ps"]`). These analyses showed that our findings were robust to choosing different prior distributions and to excluding influential studies.

@@

  # Import results from moderator analyses
  meta_cart <- read_rds("../results/results_exploratory_moderator_analyses.rds") %>% 
    group_by(kk, age, ic_direct, study_setting) %>% 
    summarize(
      r_kk = glue("$r = {f_num(median(r), digits = 2)}, [{f_num(quantile(r, 0.025), digits = 2)}, {f_num(quantile(r, 0.975), digits = 2)}]$"),
      R2 = glue("$R^2 = {f_num(median(R2) * 100, digits = 0)}\\%$")
    ) %>% 
    ungroup()

@@

Third, we used meta-regression trees to discover interactions between moderator variables that best explained heterogeneity in effect sizes [@li_meta-cart_2017; @li_multiple_2020]. Figure 6 shows the resulting meta-regression model, which explained more variance across samples than any individual moderator (`r with(meta_cart, unique(R2))`). We found that intergroup contact was associated with less perceived injustice only in studies that focused on adults and that measured intergroup contact directly. Among these studies, this association was stronger in settings in which the groups’ relative inequality stemmed from short-term migration or colonization (`r with(meta_cart, r_kk[4])`) than in other settings (`r with(meta_cart, r_kk[3])`).

@@

Developing methods to detect and correct for publication bias and other meta-biases is an active area of research, with no single method outperforming all others [@carter_correcting_2019]. Following Carter et al.’s recommendations, we compared results of several methods to adjust meta-analytic estimates for publication bias: the PET-PEESE estimator [@stanley_meta-regression_2014]; the three-parameter selection model [3PSM; @vevea_general_1995]; and the *p*-uniform* estimator [@van_aert_correcting_2018]. In addition, we ran a subgroup analysis that included only studies that were not published or not conducted with the intention to examine the effects of intergroup contact on any of the outcomes—and that, therefore, can be assumed to have not been affected by publication bias. As these methods do not use Bayesian statistics, we compared adjusted estimates to unadjusted estimates from a random-effects meta-analysis estimated with restricted maximum likelihood estimation using the *metafor* package [@viechtbauer_conducting_2010].

@@

## Alternative Predictors

>>>>

---
title: "Meta-Analysis of the 'Ironic' Effects of Intergroup Contact"
header: "'IRONIC' EFFECTS META-ANALYSIS"
keywords: intergroup contact, meta-analysis, perceived discrimination, collective action, policy support
bibliography: references.bib
csl: apa.csl
nocite: | 
  @19, @45, @46, @91, @93, @244, @284, @286, @303, @322, @324, @325, @336, @401, @414, @423, @480, @609, @703, @733, @783, @789, @803, @813, @823, @829, @830, @856, @893, @894, @933, @956, @1020, @1042, @1045, @1114, @1163, @1221, @1248, @1376, @1386, @1395, @1407, @1411, @1412, @1494, @1525, @1549, @1576, @1614, @1658, @1695, @1743, @1762, @1807, @1850, @1891, @1914, @1949, @1966, @1993, @2001, @2023, @2033, @2075, @2257, @2309, @2333, @2341, @2375, @2380, @2381, @2382, @2383, @2385, @2391, @2392, @2394, @2395, @2396, @2397, @2398, @2399, @3000, @3005, @3051, @3054, @3060, @3134, @3196, @3201, @3205, @4000, @4001, @4002, @4003, @4004, @4005, @bracegirdle_disentangling_2022, @sengupta_protect_unpublished
abstract: 
  Growing evidence suggests that intergroup contact, psychology’s most-researched paradigm for reducing prejudice, has the ‘ironic’ effect of reducing support for social change in disadvantaged groups. We conducted a preregistered meta-analytic test of this effect across `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(dplyr::n_distinct(id)))` studies with `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(length(sample)))` samples of `r with(dplyr::distinct(readRDS("../data/dl.rds"), sample, id, n), numform::f_comma(sum(n)))` disadvantaged-group members. As predicted, intergroup contact was, on average, associated with less perceived injustice (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "pi"]), digits = 2), "$"))`), collective action (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ca"]), digits = 2), "$"))`), and support for reparative policies (`r with(readRDS("../results/results_preregistered_analyses.rds"), paste0("$r = ", numform::f_num(median(r_mean[x_var == "ic" & y_var == "ps"]), digits = 2), "$"))`). However, these associations were small, variable, and consistent with alternative explanations. Across outcomes, `r with(readRDS("../results/results_for_abstract.rds"), paste0(round(min(p_pred)*100), "--", round(max(p_pred)*100), "\\%"))` of studies found *positive* associations with intergroup contact. Moderator analyses explained about a third of the between-sample variance, showing that, at least for perceived injustice, associations with intergroup contact were most consistently negative in studies that measured direct, qualitatively positive contact among adults. We also found evidence for an alternative explanation for the apparent ‘ironic’ effects of intergroup contact as, after controlling for the positive association of negative contact with support for social change, positive contact was no longer associated with any of the outcomes. We close by discussing strengths and limitations of the available evidence and by highlighting open questions about the relationship between intergroup contact and support for social change in disadvantaged groups.
output:
  pdf_document:
    latex_engine: pdflatex
    keep_tex: yes
    template: apa-template.tex
---

@@

Collective action is any action directed at improving the conditions of the disadvantaged ingroup [@wright_responding_1990]. Collective action is crucial to achieving social change because advantaged-group members rarely give up their advantages without sustained pressure from disadvantaged-group members [@blumer_race_1958; for a review, see @dixon_beyond_2012]. Researchers have argued that intergroup contact could discourage disadvantaged-group members from engaging in collective action by diminishing their perceived discrimination [@1743], by reducing the perceived threat from the advantaged outgroup [@1248], by blurring boundaries between the disadvantaged ingroup and the advantaged outgroup [@813], by reducing negative attitudes toward the advantaged outgroup [@wright_struggle_2009], and by quelling their anger  [@2309; @2375]. To the extent that it discourages collective action, intergroup contact reduces support for social change.

@@

As a result of two recent developments, there is now an emerging consensus that, in line with the former argument, intergroup contact has the 'ironic' effect of undermining support for social change in disadvantaged groups. First, a large-scale preregistered multi-country study provided the most comprehensive evidence to date that contact with advantaged groups correlates with less support for social change in disadvantaged groups [@3054]. Second, several narrative reviews of the growing evidence for the 'ironic' effects of intergroup contact have been published [@hassler_intergroup_2021; @macinnis_extending_2019; @mckeown_contact_2017]. Narrative reviews do not, however, include all relevant studies and do not synthesize the, sometimes conflicting, findings of different studies. We present the first systematic review and meta-analysis of the relationship between intergroup contact and support for social change in disadvantaged groups. To test the hypothesized 'ironic' effects of intergroup contact, we evaluate the evidence that contact with advantaged-group members reduces disadvantaged-group members' perceived injustice, collective action, and support for reparative policies.

Our first objective was to systematically review the available evidence and to evaluate its strengths and limitations. Our second objective was to synthesize the available evidence to establish the direction and magnitude of the relationships between intergroup contact and the three outcomes. Doing so allowed us to determine whether the available evidence supports a negative or positive direct relationship between contact and support for social change and to estimate the strength of this relationship. Our third objective was to estimate and explain the variability of this relationship across study settings, designs, and other potential moderators. For example, we examined whether any effects were confined to Western, Educated, Industrialized, Rich, Democratic (WEIRD) countries, the cultural context of most psychological research [@henrich_weirdest_2010]. Our fourth objective was to assess how robust the available evidence was to publication bias. Our final objective was to explore the available evidence for negative contact and ingroup contact as alternative explanations of the apparent 'ironic' effects of intergroup contact.

In pursuing these objectives, we focus on the direct relationships of intergroup contact with perceived injustice, collective action, and support for reparative policies in disadvantaged groups. We do not consider indirect and interaction effects or the effects of imagined, extended, or parasocial contact. Further, while the argument for the 'ironic' effects of intergroup contact is part of a broader critique of the prejudice reduction model of social change [for a review, see @dixon_beyond_2012], our meta-analysis tests the narrower hypothesis whether intergroup contact reduces support for social change in disadvantaged groups.

@@

We prepared a protocol in accordance with the PRISMA-P statement [@moher_preferred_2015] and preregistered it on the Open Science Framework.^[[https://osf.io/ryuev](https://osf.io/ryuev)] Preregistration precludes undisclosed flexibility in study selection, outcome selection, and other decisions that influence effect size estimates—and thus prevents confirmation bias from determining the conclusions of a meta-analysis [@lakens_reproducibility_2016]. We report deviations from our preregistered protocol at the end of the Method section. We make data, analysis scripts, and the fully reproducible manuscript available online.^[[https://osf.io/w5tqv](https://osf.io/w5tqv)]

@@

As preregistered, we searched titles, abstract, and keywords for relevant terms in four electronic databases. We searched for relevant articles in the *Scopus* and *PsycINFO* databases. We searched for gray literature in the *ProQuest Dissertations and Theses* database. We used similar non-exclusive search terms for all databases (see SOM). We searched databases on April 1, 2019 and again on April 1, 2020. We exported records and relevant metadata from each database. We removed duplicates using the *revtools* R package [@westgate_revtools:_2019].

@@

### Outcome Selection

@@

***Study Setting.*** For each sample, we recorded in what country (or countries) the data was collected ($\kappa = .97$) and what the disadvantaged ingroup and the advantaged outgroup were. We categorized each sample’s setting according to whether the source of the groups’ relative inequality was long-term migration (e.g., Asian Americans), short-term  migration (e.g., international students), slavery (e.g., Black Americans), colonization (e.g., Māori, Black South Africans), religion, caste, sexual/gender identity, or another distinction ($\kappa = .85$; see SOM for details).

***Study Design.*** We categorized each study as either observational and cross-sectional, observational and longitudinal, quasi-experimental, experimental, an intervention, or other ($\kappa = .85$).^[While the term ‘observational study’ is sometimes used to refer to the naturalistic observation of behavior, we use it in its broader sense, that is, to refer to studies that do not involve random assignment into experimental conditions [@cochran_planning_1965; @everitt_observational_2005].] We categorized each sample as either a student convenience sample, a non-student convenience sample, a probability or representative sample, or another kind of sample ($\kappa = .73$). We categorized the age group(s) in each sample as children ($\leq 12$ years), adolescents (13–18 years), or adults ($\geq 18$ years; $\kappa = .64$).

@@

***Predictor Variables.*** For each measure of intergroup contact, we coded whether it assessed contact with the specific advantaged outgroup of interest directly—or whether it assessed contact indirectly by, for example, asking what proportion of someone’s friends were not from the participants’ ingroup or what proportion of residents in someone’s neighborhood were from the relevant outgroup ($\kappa = .82$). In addition, we coded whether each predictor variable assessed direct, qualitatively positive intergroup contact by, for example, measuring contact quality or cross-group friendship with the advantaged outgroup($\kappa = .82$).^[We thank John Dixon for his suggestion to consider the distinction between contact quantity and quality in our analyses. As we added this moderator during peer review and as doing so changed the results of our moderator analyses, we report our original findings in the SOM.]

@@

We estimated effect sizes with Bayesian random-effects meta-analysis models, using the *RStan* package [@stan_development_team_rstan:_2020] for the R programming language [@r_version_2021], which modeled the *z*-transformed correlation coefficients with a normal likelihood function: \begin{align*} z_{ij} &\sim \text{Normal}(\theta_{ij}, \sigma_{ij}) \\ \theta_{ij} &= \begin{cases} \mu + \beta_j\tau_J & \text{if } I_j = 1 \\ \mu + \beta_j\tau_J + \beta_i\tau_I & \text{if } I_j > 1 \end{cases} \\ \end{align*} where $z_{ij}$ is the observed effect size in sample $i$ of study $j$, $\sigma_{ij}$ is the sample standard error, and $\theta_{ij}$ is the estimated effect size. We estimated $z_{ij}$ as a function of the mean effect size $\mu$ and of two varying (random) intercepts, $\beta_i$ and $\beta_j$, with the corresponding standard deviations, $\tau_I$ and $\tau_J$. We used the non-centered parameterization to model the random effects. For studies that contained only one sample ($I_j = 1$), we only included $\beta_j\tau_J$, the study-wise deviation from the mean effect size $\mu$. For studies that contained more than one sample ($I_j > 1$), we also estimated $\beta_i\tau_I$, the sample-wise deviation from the study-specific effect size.

@@

Bayesian inference involves choosing a likelihood function and prior distributions. A likelihood function links the observed data to one or more model parameters and states how likely the observed data are given different values of said model parameters. Prior distributions state how plausible different values of said model parameters are before considering the observed data. Bayesian inference applies Bayes’ theorem to update prior distributions in light of the observed data to produce posterior distributions. In contrast to *p*-values and confidence intervals, the resulting posterior distributions have a straightforward interpretation as stating how plausible different values of the model parameters are given the observed data.^[For an accessible introduction, see @mcelreath_statistical_2020.] We report point estimates, based on the median of posterior samples, and uncertainty intervals, based on the quantiles of posterior samples, that enclose the 95% most plausible estimates. In addition, we report the posterior probability, based on the proportion of posterior samples below zero, that the mean effect size is negative.

@@

## Deviations From the Preregistration

Some aspects of the preregistered protocol proved unrealistic, impractical, or underspecified. Therefore, we deviated from the preregistration in the following ways:

1. We preregistered that we would update our search every four months until we would submit the manuscript. This proved unrealistic as the study selection and data collection process for the new studies took almost as long. Instead, we concluded our search of electronic databases on April 1, 2020—that is, before we started analyzing and documenting our findings.

2. We preregistered that we would randomly select 100 records to be screened by both coders to calculate inter-rater agreement. We did not specify, however, what we would do if inter-rater agreement was less than acceptable. We decided to refine our coding strategy over three samples of 100 records until we achieved acceptable agreement.

3. We preregistered that we would use *Google Scholar* to find records citing eligible studies. This proved impractical as *Google Scholar* does not allow the electronic export of citing records. Instead, we used the *Scopus* citation database.

4. We preregistered that we would attempt to contact the authors of *all* papers with missing effect sizes. Instead, we decided to not contact authors for studies published before 2000 as we considered it unlikely that the authors still had access to the data.

5. We preregistered that we would use a Bayesian two-level random-effects meta-analysis model for all preregistered analyses. For the association between policy support and collective action, however, we had only few studies and found that the two-level model did not result in a reliable posterior distribution. Instead, we used a Bayesian one-level random-effects model to estimate that association.

@@

  # Leave out most influential studies
  d_loo <- read_rds("../results/results_robustness_checks_loo.rds") %>% 
    group_by(y_var, id) %>%
    median_qi(e_loo) %>% 
    group_by(y_var) %>% 
    top_n(n = 1, wt = abs(e_loo)) %>% 
    mutate(
      across(c(e_loo, .lower, .upper), ~f_num(., digits = 2)),
      text = paste0("$\\Delta r = ", e_loo, ", [", .lower, ", ", .upper, "]$")
    )

@@

As preregistered, we conducted two kinds of robustness checks. First, we assessed to what extent our findings were sensitive to choosing narrower, $\mu \sim \text{Normal}(0, 0.1)$, or wider, $\mu \sim \text{Normal}(0, 1)$, prior distributions. Choosing narrower or wider prior distribution did not affect mean effect size estimates for perceived injustice (`r d_priors$text[d_priors$y_var == "pi" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "pi" & d_priors$prior == "N(0, 1.0)"]`), collective action (`r d_priors$text[d_priors$y_var == "ca" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "ca" & d_priors$prior == "N(0, 1.0)"]`), and policy support (`r d_priors$text[d_priors$y_var == "ps" & d_priors$prior == "N(0, 0.1)"]` and `r d_priors$text[d_priors$y_var == "ps" & d_priors$prior == "N(0, 1.0)"]`). Second, we assessed to what extent our findings were sensitive to including or excluding influential studies by repeating the preregistered analyses $J$ times while leaving out one of $J$ studies each time and by calculating the mean absolute difference (*MAD*) for the estimated mean effect size across left-out studies. For perceived injustice (`r mad_loo$text[mad_loo$y_var == "pi"]`), collective action (`r mad_loo$text[mad_loo$y_var == "ca"]`), and policy support (`r mad_loo$text[mad_loo$y_var == "ps"]`), the *MAD* was small. Leaving out the most influential study, for example, did not change estimates of the mean effect sizes for the three outcomes (`r d_loo$text[d_loo$y_var == "pi"]`; `r d_loo$text[d_loo$y_var == "ca"]`; `r d_loo$text[d_loo$y_var == "ps"]`). These analyses showed that our findings were robust to choosing different prior distributions and to excluding influential studies.

@@

  # Import results from moderator analyses
  meta_cart <- read_rds("../results/results_exploratory_moderator_analyses.rds") %>% 
    group_by(kk, ic_quality, age, study_setting) %>% 
    summarize(
      r_kk = glue("$r = {f_num(median(r), digits = 2)}, [{f_num(quantile(r, 0.025), digits = 2)}, {f_num(quantile(r, 0.975), digits = 2)}]$"),
      p_kk = mean(r < 0),
      R2 = glue("$R^2 = {f_num(median(R2) * 100, digits = 0)}\\%$")
    ) %>% 
    mutate(
      p_kk = case_when(
        p_kk < .001 ~ "$\\Pr (r < 0) < 0.1\\%$",
        p_kk > .999 ~ "$\\Pr (r < 0) > 99.9\\%$",
        TRUE ~ paste0("$\\Pr (r < 0) = ", f_num(p_kk * 100, digits = 1), "\\%$")
      )
    ) %>% 
    ungroup()

@@

Third, we used meta-regression trees to discover interactions between moderator variables that best explained heterogeneity in effect sizes [@li_meta-cart_2017; @li_multiple_2020]. Figure 6 shows the resulting meta-regression model, which explained more variance across samples than any individual moderator (`r with(meta_cart, unique(R2))`). We found that intergroup contact was associated with less perceived injustice only in studies that measured direct, qualitatively positive contact among adults (`r with(meta_cart, r_kk[3])`) or in the few studies that did not measure direct, qualitatively positive contact but in which the groups’ relative inequality stemmed from short-term migration (`r with(meta_cart, r_kk[1])`).

@@

Developing methods to detect and correct for publication bias and other meta-biases is an active area of research, with no single method outperforming all others [@carter_correcting_2019]. Following Carter et al.’s recommendations, we compared results of several methods to adjust meta-analytic estimates for publication bias: the PET-PEESE estimator [@stanley_meta-regression_2014] adjusts meta-analytic estimates by correcting for the greater risk of smaller studies to not be published; the *p*-uniform* estimator [@van_aert_correcting_2018] adjusts meta-analytic estimates by correcting for the greater risk of non-significant effects to not be published; and the three-parameter selection model [3PSM, @vevea_general_1995] adjusts meta-analytic estimates for publication bias by estimating both the probability of a non-significant effect being published as well as the true average effect and between-study heterogeneity underlying the observed effect sizes.^[For an accessible introduction to meta bias correction, see @harrer_doing_2021.]

In addition, we ran a subgroup analysis that included only studies that were not published or not conducted with the intention to examine the effects of intergroup contact on any of the outcomes—and that, therefore, can be assumed to have not been affected by publication bias. As the methods we used do not use Bayesian statistics, we compared adjusted estimates to unadjusted estimates from a random-effects meta-analysis estimated with restricted maximum likelihood estimation using the *metafor* R package [@viechtbauer_conducting_2010].

@@

## Additional Analyses

```{=latex}
\begin{figure*}[t!]
\centering
\caption{Posterior distributions for studies that measured direct, qualitatively positive contact among adults}
\includegraphics[scale=1]{../figures/figure-8}
\caption*{\textit{Note.} Black outlines show the posterior distributions from the original, preregistered analyses. \textbf{A} Posterior distributions for the mean correlation coefficients, highlighting the proportion of posterior samples for which $r_\text{mean} < 0$. \textbf{B} Posterior predictive distributions of study-wise correlation coefficients, highlighting the 80\% most common effect sizes, with point estimates for the correlation coefficients for all studies in the sample.}
\label{fig:f8}
\end{figure*}
```

  # Import results from analyses
  pre_results <- read_rds("../results/results_preregistered_analyses.rds")
  add_results <- read_rds("../results/results_additional_analyses.rds")
  add_r_pred_jj <- read_rds("../results/r_additional_analyses_pred_jj.rds")

  # Merge results
  results <- add_results %>% 
    select(.chain:.draw, x_var, y_var, I, J, N, r_add = r_mean) %>% 
    left_join(
      pre_results %>% 
        select(.chain:.draw, x_var, y_var, r_pre = r_mean),
      by = c(".chain", ".iteration", ".draw", "x_var", "y_var")
    ) %>% 
    group_by(y_var) %>%
    summarize(
      I = f_comma(unique(I)),
      J = f_comma(unique(J)),
      N = f_comma(unique(N)),
      r_median = median(r_add),
      r_lower = quantile(r_add, 0.025),
      r_upper = quantile(r_add, 0.975),
      d_median = median(r_add - r_pre),
      d_lower = quantile(r_add - r_pre, 0.025),
      d_upper = quantile(r_add - r_pre, 0.975),
      p_below = mean(r_add < 0)
    ) %>% 
    left_join(
      add_r_pred_jj %>% 
        filter(x_var == "ic") %>% 
        group_by(y_var) %>% 
        summarize(p_pred = mean(r_pred > 0)),
      by = "y_var"
    ) %>% 
    mutate(
      across(
        c(starts_with("r_"), starts_with("d_")), 
        ~f_num(., digits = 2)
      ),
      r_text = paste0("$r = ", r_median, ", [", r_lower, ", ", r_upper, "]$"),
      d_text = paste0("$\\Delta r = ", d_median, ", [", d_lower, ", ", d_upper, "]$"),
      p_below = case_when(
        p_below < .001 ~ "$\\Pr (r < 0) < 0.1\\%$",
        p_below > .999 ~ "$\\Pr (r < 0) > 99.9\\%$",
        TRUE ~ paste0("$\\Pr (r < 0) = ", f_num(p_below * 100, digits = 1), "\\%$")
      ),
      p_pred = case_when(
        p_pred < .001 ~ "$< 0.1\\%$",
        p_pred > .999 ~ "$> 99.9\\%$",
        TRUE ~ paste0("$", f_num(p_pred * 100, digits = 1), "\\%$")
      )
    )

Our moderator analyses suggested that, for perceived injustice, the evidence for the 'ironic' effect of intergroup contact was stronger in studies that measured direct, qualitatively positive contact among adults than in most other studies. Operationalizing intergroup contact as direct, qualitatively positive contact among adults is closer to the original argument for the 'ironic' effects of intergroup contact and, therefore, tests a stronger case for the hypothesized relationships. To test this stronger case, we repeated our original, preregistered analyses for studies that measured direct, qualitatively positive contact among adults.
